High-Dimensional Regularized Discriminant Analysis

### 2013-05-11 reports/paper-hdrda.Rnw
Improved the abstract.

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, because the number of gene expression levels measured is
typically much greater than the sample size, the analysis of microarray data
often presents difficult statistical and computational challenges. Here, we
propose a new regularized discriminant analysis classifier that incorporates
dimension reduction and covariance-matrix shrinkage to enable a computationally
efficient classifier and model-selection procedure. We demonstrate with four
high-dimensional microarray data sets that our proposed classifier is often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ towards $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

>>>>

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, because the number of gene expression levels measured is
typically much greater than the sample size, the analysis of microarray data can
be difficult because classic classifiers are often ill-posed in this setting and
require a large number of calculations. To address these challenges, we propose
a new regularized discriminant analysis classifier that incorporates dimension
reduction and covariance-matrix shrinkage to enable an efficacious,
computationally efficient method. We demonstrate with four high-dimensional
microarray data sets that our proposed classifier is often superior to several
recently proposed sparse and regularized classifiers in terms of classification
accuracy. Finally, we provide an implementation of our proposed classifier in
the {\tt sparsediscrim} R package, available on CRAN.

@@

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.
